---
title: "Rmarkdown basics"
author: "Cylcia Bolibaugh"
output: 
    html_document:
        code_download: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r lib, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
# devtools::install_github("debruine/faux") 
# library(faux)
library(academicWriteR)
```
Here is some code and output from last week's simulations. We'll use it this week to see how we can embed figures, set chunk options, and use stats inline.

## Simulate independent samples

We started with a simple between subjects design, where the independent (grouping) variable has two levels, and the dependent variable is from a normal distribution. To simulate this dataset, we need to know

* How many subjects are in each condition?
* What are the score means (for each level of the independent variable)?
* What are the score variances?

### Parameters {#ind-params}

We started our simulation script by setting parameters for these values. Here these are the population values from which we'll sample.

_What happens if we change the chunk option to "echo = FALSE"?_

```{r ind-vars, echo = TRUE}
#simulate normally distributed data for an independent samples t-test
muA <- 50 #what's the population mean for condition A?
muB <- 55 #what's the population mean for condition B?

n1 <- 100 #how many participants group A?
n2 <- 100 #how many participants group B?

sigma1 <- 20 #population standard deviation of condition A?
sigma2 <- 20 #population standard deviation of condition B?
```

### Scores

You can generate the scores using the `rnorm()` function (notice the code is  shown here for demo purposes), then plot the distribution of scores for each group.

```{r ind-dat, echo=TRUE}
A_scores <- rnorm(n1,muA,sigma1) #simulate a vector of 100 deviates from normal distribution
B_scores <- rnorm(n2,muB,sigma2) #simulate a vector of 100 deviates from normal distribution
```
```{r histA, fig.cap = "Figure 1: Distribution of scores from Group A"}
hist(A_scores)
```

```{r histB, fig.cap = "Figure 2: Distribution of scores from Group B"}
hist(B_scores)
```
###  Analyses   

You can then run your inferential test on your new data.
```{r t-test1}
t.test(A_scores,B_scores)
```
If we want to report the results from the statistical test in an inline report (i.e. a sentence), we can assign it to an object and extract the parameters we need in text:
```{r t-test2}
mod1 <- broom::tidy(t.test(A_scores,B_scores))
```

    If we include the following inline text, we can include results directly without copying or pasting:    
    
A `` `r
mod1$method` `` indicated that participants in condition A (*M* = `` `r
mean(A_scores)` ``, *SD* = `` `r
sd(A_scores)` ``) did [not] score significantly lower than participants in condition B (*M* = `` `r
mean(B_scores)` ``, 
*SD* = `` `r 
sd(B_scores)` ``), 
*t*(`` `r
mod1$parameter` ``) = 
`` `r
mod1$statistic` ``, 
*p* = `` `r
mod1$p.value` ``.

This will print out as follows:  

> A `r mod1$method` indicated that participants in condition A (*M* = `r mean(A_scores)`, *SD* = `r sd(A_scores)`) did [not] score significantly lower than participants in condition B (*M* =`r mean(B_scores)`, *SD* = `r sd(B_scores)`), *t*(`r mod1$parameter`) = `r mod1$statistic`, *p* = `r mod1$p.value`.



 Unfortunately, the number of digits isn't helpful. The `round` function can help:
 
> A `r mod1$method` indicated that participants in condition A (*M* = `r round(mean(A_scores),2)`, *SD* = `r round(sd(A_scores),2)`) did [not] score significantly lower than participants in condition B (*M* = `r round(mean(B_scores),2)`, *SD* = `r round(sd(B_scores),2)`), *t*(`r round(mod1$parameter,2)`) = `r round(mod1$statistic,2)`, *p* = `r round(mod1$p.value,2)`.
 
### APA formatting
Even with the help of the round function, there are still a number of formatting issues which would need to be fixed if we were drafting an APA style report or manuscript. Luckily the `Papaja` [package](https://crsh.github.io/papaja_man/index.html) has automated many of these functions. 

You can try it out by opening a new Rmd document, installing Papaja from the GitHub repository:
install.packages("devtools")  
devtools::install_github("crsh/papaja"),  
and choosing the template.
